---
title: "ExploratoryAnalysis"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up the packages that are needed here

```{r}
library(SIBER)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
```

Set up the asthetics

```{r}
# Define a unified palette for communities and groups
# Define updated colors with a more visible option for Yellow Pond Lily
unified_colors <- c(
  "Algae" = "red", 
  "Amphipod" = "blue", 
  "Beetle" = "green", 
  "Bladderwort" = "purple", 
  "Bogbean" = "orange", 
  "Burreed" = "pink", 
  "Caddisfly" = "cyan", 
  "Clams" = "brown", 
  "Cotton Grass" = "darkgreen",
  "Damselfly" = "gold", 
  "Dragonfly" = "black", 
  "Elodea" = "grey", 
  "Fly" = "navy", 
  "Horsetail" = "lightblue", 
  "Leech" = "magenta",
  "Mare's Tail" = "darkred", 
  "Milfoil" = "yellowgreen", 
  "Mite" = "lightpink", 
  "Peat Moss" = "turquoise", 
  "Pondweed" = "violet", 
  "Purple Marshlocks" = "darkblue", 
  "Reed Grass" = "darkorange", 
  "Snail" = "khaki", 
  "Water Boatman" = "plum", 
  "Willow" = "coral", 
  "Yellow Pond Lily" = "#1F77B4"  # Bright gold for better visibility
)


# Define 27 shapes for community and 17 for macrophytes
unified_shapes <- c(
  "Algae" = 0, "Amphipod" = 1, "Beetle" = 2, "Bladderwort" = 3, 
  "Bogbean" = 4, "Burreed" = 5, "Caddisfly" = 6, "Clams" = 7, 
  "Cotton Grass" = 8, "Damselfly" = 9, "Dragonfly" = 10, "Elodea" = 11, 
  "Fly" = 12, "Horsetail" = 13, "Leech" = 14, "Mare's Tail" = 15, 
  "Milfoil" = 16, "Mite" = 17, "Peat Moss" = 18, "Pondweed" = 19, 
  "Purple Marshlocks" = 20, "Reed Grass" = 21, "Snail" = 22, 
  "Water Boatman" = 23, "Willow" = 24, "Yellow Pond Lily" = 25
)

```

# Read in the data file -- for the total SI data (inverts plus macrophytes)

```{r}
# import the data. Replace this line with a read.csv() or similar call
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_SIData.csv",
                   header=TRUE)
demo_data <- mydata %>% mutate(group = factor(group), 
                               community = factor(community),
                               d13C = iso1, 
                               d15N = iso2,
                               .keep = "unused") 
TestData <- demo_data %>%
  mutate(status = ifelse(group == "Swanny", "Elodea",
                         ifelse(group == "WestCanneryWest", "Elodea",
                                ifelse(group == "Wigeon", "Elodea",
                                       ifelse(group == "ThePinch", "Native",
                                              ifelse(group == "EyakSouth", "Native",
                                                     ifelse(group == "Strawberry", "Native", "blank")))))))
TestData2 <- TestData %>%
  mutate(pair = ifelse(group == "Swanny", "Alaganik",
                         ifelse(group == "WestCanneryWest", "Eyak",
                                ifelse(group == "Wigeon", "Road",
                                       ifelse(group == "ThePinch", "Alaganik",
                                              ifelse(group == "EyakSouth", "Eyak",
                                                     ifelse(group == "Strawberry", "Road", "blank")))))))

# appears this worked to add the status column
demo_data <- TestData2
```

# Setting up the biplot here

```{r}
# when plotting colors and shapes, we need to tell ggplot that these are to 
# be treated as categorical factor type data, and not numeric.

#first I need to set up the shapes since ggplot will not make a figure for it
# if it has more than 6 community members
colors <- c("blue", "red")


first.plot <- ggplot(data = demo_data, 
                     aes(x = d13C, 
                         y = d15N)) + 
  geom_point(aes(shape = community, color = community), size = 2) +
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(#plot.title = element_text(hjust = 0.5), 
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 13),
    strip.text.x = element_text(size = 20),
    strip.text.y = element_text(size = 20))

# And print our plot to screen
print(first.plot)

bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18))

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)
```

Making this have red and blue lables for Status (elodea and native)

```{r}
g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Biplot_all.png", plot = g, height = 6, width = 12, dpi = 300)
```

# Plot to put into inkscape

```{r}
library(ggplot2)
library(gridExtra)  # For combining the legend and the plot

# First, create the plot with the legend
first.plot <- ggplot(data = demo_data, 
                     aes(x = d13C, 
                         y = d15N)) + 
  geom_point(aes(shape = community, color = community), size = 2) +
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic()+
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 13),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 20))

# Remove the legend from the main plot
first.plot_no_legend <- first.plot + theme(legend.position = "none")

# Print the main plot (without the legend)
print(first.plot_no_legend)


bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18),
        legend.position = "none")

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)

g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Biplot_all_PUB.png", plot = g, height = 6, width = 12, dpi = 300)

# Extract the legend from the plot
get_legend <- function(plot) {
  tmp <- ggplot_gtable(ggplot_build(plot))
  leg <- tmp$grobs[[which(sapply(tmp$grobs, function(x) x$name) == "guide-box")]]
  return(leg)
}

legend_plot <- get_legend(first.plot)

# Now create a plot with just the legend
legend_only_plot <- grid.arrange(legend_plot, ncol = 1)

# Save the legend plot separately
ggsave("Analysis_Fall2024/Plots/new/CRD_Legend_only.png", plot = legend_only_plot, height = 3, width = 6, dpi = 300)


```


Make the same plot of only the macrophytes

# Read in the data file -- for the total SI data (macrophytes)

```{r}
# import the data. Replace this line with a read.csv() or similar call
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_SI_Macrophyte.csv",
                   header=TRUE)
demo_data <- mydata %>% mutate(group = factor(group), 
                               community = factor(community),
                               d13C = iso1, 
                               d15N = iso2,
                               .keep = "unused") 
TestData <- demo_data %>%
  mutate(status = ifelse(community == "SWN", "Elodea",
                         ifelse(community == "WCW", "Elodea",
                                ifelse(community == "WGN", "Elodea",
                                       ifelse(community == "PCH", "Native",
                                              ifelse(community == "EYS", "Native",
                                                     ifelse(community == "STB", "Native", "blank")))))))
TestData2 <- TestData %>%
  mutate(pair = ifelse(community == "SWN", "Alaganik",
                         ifelse(community == "WCW", "Eyak",
                                ifelse(community == "WGN", "Road",
                                       ifelse(community == "PCH", "Alaganik",
                                              ifelse(community == "EYS", "Eyak",
                                                     ifelse(community == "STB", "Road", "blank")))))))

# appears this worked to add the status column
demo_data <- TestData2
```

# Setting up the biplot here

```{r}
# when plotting colors and shapes, we need to tell ggplot that these are to 
# be treated as categorical factor type data, and not numeric.

#first I need to set up the shapes since ggplot will not make a figure for it
# if it has more than 6 community members
colors <- c("blue", "red")


first.plot <- ggplot(data = demo_data, 
                     aes(x = d13C, 
                         y = d15N)) + 
  geom_point(aes(shape = group, color = group), size = 2) +
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(#plot.title = element_text(hjust = 0.5), 
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 13),
    strip.text.x = element_text(size = 20),
    strip.text.y = element_text(size = 20))

# And print our plot to screen
print(first.plot)

bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18))

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)
```

Making this have red and blue lables for Status (elodea and native)

```{r}
library(grid)
g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Macrophyte_Biplot_all.png", plot = g, height = 6, width = 12, dpi = 300)
```

# Plot to put into inkscape

```{r}
library(ggplot2)
library(gridExtra)  # For combining the legend and the plot

# First, create the plot with the legend
first.plot <- ggplot(data = demo_data, 
                     aes(x = d13C, 
                         y = d15N)) + 
  geom_point(aes(shape = group, color = group), size = 2) +
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 13),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 20))

# Remove the legend from the main plot
first.plot_no_legend <- first.plot + theme(legend.position = "none")

# Print the main plot (without the legend)
print(first.plot_no_legend)


bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18),
        legend.position = "none")

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)

g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Macrophyte_Biplot_all_PUB.png", plot = g, height = 6, width = 12, dpi = 300)
```


Making a plot with the Elodea samples for WCW compared to the Elodea samples from SWN and WGN

```{r}
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_MacrophyteData_SIBER.csv", header=TRUE)

elodea <- mydata %>%
  filter (group == "Elodea")

# Define shapes and colors for each community
shapes <- c(16, 17, 18)  # Circle, Triangle, Square (for 3 communities)
colors <- c("blue", "red", "green")  # Assign colors for each community

# Create the plot
plot <- ggplot(data = elodea, 
               aes(x = iso1, y = iso2)) + 
  geom_point(aes(shape = community, color = community), size = 3) +
  scale_shape_manual(values = shapes) +  # Map community to shape
  scale_color_manual(values = colors) +  # Map community to color
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 13)
  ) +
  guides(shape = guide_legend(title = "Community"), 
         color = guide_legend(title = "Community"))

# Show the plot
plot

```

Creating the graph of the Elodea imputed values

```{r}
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_MacrophyteData_SIBER.csv", header=TRUE)
```

Plotting these

```{r}
#first select only Elodea data
elodea <- mydata %>%
  filter(group == "Elodea")

Elodea_plot <- ggplot(data = elodea, aes(x = iso1, y = iso2)) + 
  geom_point(aes(color = community, shape = community), size = 3) + # Map both color and shape to 'community'
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 15)
  ) +
  guides(
    shape = guide_legend(title = "Community"),  # Single legend for 'Community'
    color = "none"  # Remove the color legend
  )
Elodea_plot

ggsave("Analysis_Fall2024/Plots/new/Elodea_Biplot.png", plot = Elodea_plot, height = 6, width = 12, dpi = 300)

```



# Making a plot of the isotope data with it as a mean +/- standard deviation

Start by defining the aethetics

```{r}
# Define a unified palette for communities and groups
# Define updated colors with a more visible option for Yellow Pond Lily
unified_colors <- c(
  "Algae" = "red", 
  "Amphipod" = "blue", 
  "Beetle" = "green", 
  "Bladderwort" = "purple", 
  "Bogbean" = "orange", 
  "Burreed" = "pink", 
  "Caddisfly" = "cyan", 
  "Clams" = "brown", 
  "Cotton Grass" = "darkgreen",
  "Damselfly" = "gold", 
  "Dragonfly" = "black", 
  "Elodea" = "grey", 
  "Fly" = "navy", 
  "Horsetail" = "lightblue", 
  "Leech" = "magenta",
  "Mare's Tail" = "darkred", 
  "Milfoil" = "yellowgreen", 
  "Mite" = "lightpink", 
  "Peat Moss" = "turquoise", 
  "Pondweed" = "violet", 
  "Purple Marshlocks" = "darkblue", 
  "Reed Grass" = "darkorange", 
  "Snail" = "khaki", 
  "Water Boatman" = "plum", 
  "Willow" = "coral", 
  "Yellow Pond Lily" = "#1F77B4"  # Bright gold for better visibility
)


# Define 27 shapes for community and 17 for macrophytes
unified_shapes <- c(
  "Algae" = 0, "Amphipod" = 1, "Beetle" = 2, "Bladderwort" = 3, 
  "Bogbean" = 4, "Burreed" = 5, "Caddisfly" = 6, "Clams" = 7, 
  "Cotton Grass" = 8, "Damselfly" = 9, "Dragonfly" = 10, "Elodea" = 11, 
  "Fly" = 12, "Horsetail" = 13, "Leech" = 14, "Mare's Tail" = 15, 
  "Milfoil" = 16, "Mite" = 17, "Peat Moss" = 18, "Pondweed" = 19, 
  "Purple Marshlocks" = 20, "Reed Grass" = 21, "Snail" = 22, 
  "Water Boatman" = 23, "Willow" = 24, "Yellow Pond Lily" = 25
)

```

# Read in the data file -- for the total SI data (inverts plus macrophytes)

```{r}
# import the data. Replace this line with a read.csv() or similar call
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_SIData.csv",
                   header=TRUE)
demo_data <- mydata %>% mutate(group = factor(group), 
                               community = factor(community),
                               d13C = iso1, 
                               d15N = iso2,
                               .keep = "unused") 
TestData <- demo_data %>%
  mutate(status = ifelse(group == "Swanny", "Elodea",
                         ifelse(group == "WestCanneryWest", "Elodea",
                                ifelse(group == "Wigeon", "Elodea",
                                       ifelse(group == "ThePinch", "Native",
                                              ifelse(group == "EyakSouth", "Native",
                                                     ifelse(group == "Strawberry", "Native", "blank")))))))
TestData2 <- TestData %>%
  mutate(pair = ifelse(group == "Swanny", "Alaganik",
                         ifelse(group == "WestCanneryWest", "Eyak",
                                ifelse(group == "Wigeon", "Road",
                                       ifelse(group == "ThePinch", "Alaganik",
                                              ifelse(group == "EyakSouth", "Eyak",
                                                     ifelse(group == "Strawberry", "Road", "blank")))))))

# appears this worked to add the status column
demo_data <- TestData2
```

## Starting with the whole community

Calculate the mean +/- sd for each taxa for each pond

```{r}
library(dplyr)
library(ggplot2)

# Calculate mean and SD for each community/group combination
summary_data <- demo_data %>%
  group_by(community, group, status, pair) %>%
  summarise(
    mean_d13C = mean(d13C),
    sd_d13C = sd(d13C),
    mean_d15N = mean(d15N),
    sd_d15N = sd(d15N),
    .groups = "drop"
  )
summary_data
```

Create the panel plot

```{r}
panel_plot <- ggplot(data = summary_data, 
                     aes(x = mean_d13C, 
                         y = mean_d15N)) +
  # Error bars first (behind points)
  geom_errorbar(aes(ymin = mean_d15N - sd_d15N, ymax = mean_d15N + sd_d15N, color = community), 
                width = 0.2, size = 0.8) +
  geom_errorbarh(aes(xmin = mean_d13C - sd_d13C, xmax = mean_d13C + sd_d13C, color = community), 
                 height = 0.2, size = 0.8) +
  # Points second (on top of error bars)
  geom_point(aes(shape = community, color = community), size = 4) +  # Increase size of points
  guides(shape = guide_legend(title = "Community"), color = guide_legend(title = "Community")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 13),
    strip.text.x = element_text(size = 20),
    strip.text.y = element_text(size = 20)
  )
print(panel_plot)

```

Update to B&W theme

```{r}
# Add black-and-white theme
bw_panel_plot <- panel_plot + 
  theme_bw() +
  theme(text = element_text(size = 18))

# Define lake labels
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

# Add labels
final_plot <- bw_panel_plot + 
  geom_text(
    data = dat_text,
    mapping = aes(x = -Inf, y = -Inf, label = label),
    hjust = -0.1,
    vjust = -1
  )
print(final_plot)

```

Update to have the red and blue lables

```{r}
library(grid)
g <- ggplot_gtable(ggplot_build(final_plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D", "#619CFF")  # Define red and blue fills
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Biplot.png", plot = g, height = 6, width = 12, dpi = 300)
```

# Plot for inkscape figure

```{r}
library(ggplot2)
library(gridExtra)  # For combining the legend and the plot

# First, create the plot with the legend
first.plot <- ggplot(data = summary_data, 
                     aes(x = mean_d13C, 
                         y = mean_d15N)) +
  # Error bars first (behind points)
  geom_errorbar(aes(ymin = mean_d15N - sd_d15N, ymax = mean_d15N + sd_d15N, color = community), 
                width = 0.2, size = 0.8) +
  geom_errorbarh(aes(xmin = mean_d13C - sd_d13C, xmax = mean_d13C + sd_d13C, color = community), 
                 height = 0.2, size = 0.8) +
  # Points second (on top of error bars)
  geom_point(aes(shape = community, color = community), size = 4) +  # Increase size of points
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 13),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 20))

# Remove the legend from the main plot
first.plot_no_legend <- first.plot + theme(legend.position = "none")

# Print the main plot (without the legend)
print(first.plot_no_legend)


bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18),
        legend.position = "none")

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)

g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Biplot_Sum_PUB.png", plot = g, height = 6, width = 12, dpi = 300)
```

## Repeat for the Macrophyte communities

```{r}
# import the data. Replace this line with a read.csv() or similar call
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_SI_Macrophyte.csv",
                   header=TRUE)
demo_data <- mydata %>% mutate(group = factor(group), 
                               community = factor(community),
                               d13C = iso1, 
                               d15N = iso2,
                               .keep = "unused") 
TestData <- demo_data %>%
  mutate(status = ifelse(community == "SWN", "Elodea",
                         ifelse(community == "WCW", "Elodea",
                                ifelse(community == "WGN", "Elodea",
                                       ifelse(community == "PCH", "Native",
                                              ifelse(community == "EYS", "Native",
                                                     ifelse(community == "STB", "Native", "blank")))))))
TestData2 <- TestData %>%
  mutate(pair = ifelse(community == "SWN", "Alaganik",
                         ifelse(community == "WCW", "Eyak",
                                ifelse(community == "WGN", "Road",
                                       ifelse(community == "PCH", "Alaganik",
                                              ifelse(community == "EYS", "Eyak",
                                                     ifelse(community == "STB", "Road", "blank")))))))

# appears this worked to add the status column
demo_data2 <- TestData2
```

Calculating the mean +/- sd for each of the taxa in each pond

```{r}
library(dplyr)
library(ggplot2)

summary_data2 <- demo_data2 %>%
  group_by(group, community, status, pair) %>%  # Include status and pair in the grouping
  summarise(
    mean_d13C = mean(d13C),
    sd_d13C = sd(d13C),
    mean_d15N = mean(d15N),
    sd_d15N = sd(d15N),
    .groups = "drop"
  )
summary_data2

```

Creating the panel plot

```{r}
panel_plot2 <- ggplot(data = summary_data2, 
                     aes(x = mean_d13C, 
                         y = mean_d15N)) +
  # Error bars first (appear behind points)
  geom_errorbar(aes(ymin = mean_d15N - sd_d15N, ymax = mean_d15N + sd_d15N, color = group), 
                width = 0.2, size = 0.8) +
  geom_errorbarh(aes(xmin = mean_d13C - sd_d13C, xmax = mean_d13C + sd_d13C, color = group), 
                 height = 0.2, size = 0.8) +
  # Points second (appear on top of error bars)
  geom_point(aes(shape = group, color = group), size = 4) +  # Increase size of points
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 13),
    strip.text.x = element_text(size = 20),
    strip.text.y = element_text(size = 20)
  )
print(panel_plot2)


```

```{r}
# Add black-and-white theme
bw_panel_plot2 <- panel_plot2 + 
  theme_bw() +
  theme(text = element_text(size = 18))

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

# Add labels
final_plot2 <- bw_panel_plot2 + 
  geom_text(
    data = dat_text,
    mapping = aes(x = -Inf, y = -Inf, label = label),
    hjust = -0.1,
    vjust = -1
  )
print(final_plot2)
```

Addin the red and blue labels here 

```{r}
library(grid)
g <- ggplot_gtable(ggplot_build(final_plot2))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Macrophyte_Biplot.png", plot = g, height = 6, width = 10, dpi = 300)
```

# Plot for inkscape

# Plot for inkscape figure

```{r}
library(ggplot2)
library(gridExtra)  # For combining the legend and the plot

# First, create the plot with the legend
first.plot <- ggplot(data = summary_data2, 
                     aes(x = mean_d13C, 
                         y = mean_d15N)) +
  # Error bars first (behind points)
  geom_errorbar(aes(ymin = mean_d15N - sd_d15N, ymax = mean_d15N + sd_d15N, color = group), 
                width = 0.2, size = 0.8) +
  geom_errorbarh(aes(xmin = mean_d13C - sd_d13C, xmax = mean_d13C + sd_d13C, color = group), 
                 height = 0.2, size = 0.8) +
  # Points second (on top of error bars)
  geom_point(aes(shape = group, color = group), size = 4) +  # Increase size of points
  guides(shape = guide_legend(title = "Taxa"), color = guide_legend(title = "Taxa")) +
  scale_color_manual(values = unified_colors) +
  scale_shape_manual(values = unified_shapes) +
  facet_grid(rows = vars(status), cols = vars(pair)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 13),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 20))

# Remove the legend from the main plot
first.plot_no_legend <- first.plot + theme(legend.position = "none")

# Print the main plot (without the legend)
print(first.plot_no_legend)


bw.first.plot <- first.plot + theme_bw() + 
  theme(text = element_text(size=18),
        legend.position = "none")

# and print to screen
print(bw.first.plot)

# Try to add lake labels to this
dat_text <- data.frame(
  label = c("SWN", "PCH", "WCW", "EYS", "WGN", "STB"),
  pair = c("Alaganik", "Alaganik", "Eyak", "Eyak", "Road", "Road"),
  status = c("Elodea", "Native", "Elodea", "Native", "Elodea", "Native")
)

plot <- bw.first.plot + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -0.1,
  vjust   = -1
)

print(plot)

g <- ggplot_gtable(ggplot_build(plot))
strip_right <- which(grepl('strip-r', g$layout$name))
fills <- c("#F8766D","#619CFF")
k <- 1
for (i in strip_right) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.draw(g)

ggsave("Analysis_Fall2024/Plots/new/CRD_Macrophyte_Biplot_Sum_PUB.png", plot = g, height = 6, width = 12, dpi = 300)
```
