---
title: "SIBER_Analysis"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting up the workspace

```{r}
library(SIBER, quietly = TRUE,
        verbose = FALSE,
        logical.return = FALSE)
library(viridis)
library(tibble)
```

Choose 4 colours from the viridis color palette because
We have 4 groups in each community and we want to plot them in different colours.
```{r}
palette(viridis(4))
```

Read in the data
```{r}
mydata <- read.csv("Analysis_Fall2024/DataFiles/CRD_MacrophyteData_SIBER.csv", header=TRUE)
```

Create the siber object

```{r}
siber.example <- createSiberObject(mydata)
```

Create lists of plotting arguments to be passed onwards to each of the three plotting functions.

```{r}
community.hulls.args <- list(col = "black", lty = 1, lwd = 1)
group.ellipses.args  <- list(n = 100, p.interval = 0.95, lty = 1, lwd = 2)
group.hull.args      <- list(lty = 2, col = "grey20")

# plot the raw data
par(mfrow=c(1,1))
plotSiberObject(siber.example,
                ax.pad = 2, 
                hulls = TRUE, community.hulls.args, 
                ellipses = FALSE, group.ellipses.args,
                group.hulls = FALSE, group.hull.args,
                bty = "L",
                iso.order = c(1,2),
                xlab = expression({delta}^13*C~'\u2030'),
                ylab = expression({delta}^15*N~'\u2030')
)

# add the confidence interval of the means to help locate
# the centre of each data cluster

## This section below does not work because different communities within each group
#plotGroupEllipses(siber.example, n = 100, p.interval = 0.95,
 #                 ci.mean = TRUE, lty = 1, lwd = 2)
```

# Fit the Bayesian models

Options for running jags

```{r}
parms <- list()
parms$n.iter <- 2 * 10^4   # number of iterations to run the model for
parms$n.burnin <- 1 * 10^3 # discard the first set of values
parms$n.thin <- 10     # thin the posterior by this many
parms$n.chains <- 2        # run this many chains
```

Define the priors

```{r}
priors <- list()
priors$R <- 1 * diag(2)
priors$k <- 2
priors$tau.mu <- 1.0E-3
```

Fit the ellipses which uses an Inverse Wishart prior on the covariance matrix Sigma, and a vague normal prior on the means. Fitting is via the JAGS method.

```{r}
ellipses.posterior <- siberMVN(siber.example, parms, priors)
```

Extract the posterior means

```{r}
#This section also does not work with different communities per group
 mu.post <- extractPosteriorMeans(siber.example, ellipses.posterior)
```

Calculate the corresponding distribution of layman metrics

```{r}
layman.B <- bayesianLayman(mu.post)
```

# Visualise the first community

Drop the 3rd column of the posterior which is TA using -3. 
Makes this hard to view

```{r}
siberDensityPlot(layman.B[[1]][ , -3], 
               xticklabels = colnames(layman.B[[1]][ , -3]), 
               bty="L", ylim = c(0,20))
```

Add the ML estimates (if you want). Extract the correct means from the appropriate array held within the overall array of means.

```{r}
#comm1.layman.ml <- laymanMetrics(siber.example$ML.mu[[1]][1,1,],
 #                               siber.example$ML.mu[[1]][1,2,]
#)

#again drop the 3rd entry which relates to TA
#points(1:5, comm1.layman.ml$metrics[-3], col = "red", pch = "x", lwd = 2)
```

# Visualise the second community

```{r}
#This section also does not work with different communities per group
#siberDensityPlot(layman.B[[2]][ , -3], 
 #                xticklabels = colnames(layman.B[[2]][ , -3]), 
  #               bty="L", ylim = c(0,20))
```

Add the ML estimates. (if you want) Extract the correct means from the appropriate array held within the overall array of means.

```{r}
#This section also does not work with different communities per group
#comm2.layman.ml <- laymanMetrics(siber.example$ML.mu[[2]][1,1,],
 #                                siber.example$ML.mu[[2]][1,2,]
#)
#points(1:5, comm2.layman.ml$metrics[-3], 
 #      col = "red", pch = "x", lwd = 2)
```

Alternatively, pull out TA from both and aggregate them into a single matrix using cbind() and plot them together on one graph.

Go back to a 1x1 panel plot

```{r}
par(mfrow=c(1,1))
```

Now we only plot the TA data. We could address this as either
  layman.B[[1]][, "TA"]
        or 
  layman.B[[1]][, 3]
  
```{r}
siberDensityPlot(cbind(layman.B[[1]][ , "TA"], 
                       layman.B[[2]][ , "TA"],
                       layman.B[[3]][ , "TA"], 
                       layman.B[[4]][ , "TA"],
                       layman.B[[5]][ , "TA"], 
                       layman.B[[6]][ , "TA"]),
                 xticklabels = c("EYS", "PCH","STB", "SWN","WCW", "WGN"), 
                 bty="L", ylim = c(0,60),
                 las = 1,
                 ylab = "TA - Convex Hull Area",
                 xlab = "")

TA_means <- c(mean(layman.B[[1]][ , "TA"]), mean(layman.B[[2]][ , "TA"]), mean(layman.B[[3]][ , "TA"]), 
               mean(layman.B[[4]][ , "TA"]), mean(layman.B[[5]][ , "TA"]), mean(layman.B[[6]][ , "TA"]))
TA_sds <- c(sd(layman.B[[1]][ , "TA"]), sd(layman.B[[2]][ , "TA"]), sd(layman.B[[3]][ , "TA"]), 
                sd(layman.B[[4]][ , "TA"]), sd(layman.B[[5]][ , "TA"]), sd(layman.B[[6]][ , "TA"]))
names <- c("EYS", "PCH","STB", "SWN","WCW", "WGN")
Status <- c("Native", "Native", "Native", "Elodea", "Elodea", "Elodea")

TotalArea <- tibble(names, Status, TA_means, TA_sds)

#Trying to make these plots in ggplot instead as violin plots

# Load required package
library(ggplot2)
library(dplyr)

# Combine the data into a single data frame
layman_data <- do.call(rbind, lapply(seq_along(layman.B), function(i) {
  data.frame(Group = paste0("Group_", i), TA = layman.B[[i]][, "TA"])
}))

# Define the group names for labeling
group_labels <- c("EYS", "PCH", "STB", "SWN", "WCW", "WGN")

# Calculate the 2.5th and 97.5th percentiles for each group
layman_data_filtered <- layman_data %>%
  group_by(Group) %>%
  mutate(
    lower_cutoff = quantile(TA, 0.025),
    upper_cutoff = quantile(TA, 0.975)
  ) %>%
  filter(TA >= lower_cutoff & TA <= upper_cutoff) %>%
  ungroup()

# Calculate medians for each group
median_values <- layman_data_filtered %>%
  group_by(Group) %>%
  summarize(median_TA = median(TA), .groups = "drop")

# Create the violin plot
TotalArea_plot <- ggplot(layman_data_filtered, aes(x = factor(Group, labels = group_labels), y = TA, fill = Group)) +
  geom_violin(trim = TRUE) +
  geom_point(data = median_values, aes(x = factor(Group, labels = group_labels), y = median_TA), 
             shape = 4, size = 3, color = "black", stroke = 1.2) + # Add X for median
  ylab("Total Area") +
  xlab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 14),     # Increase size of axis tick labels
    axis.title = element_text(size = 20),   # Increase size of axis titles
    axis.text.x = element_text(size = 16),  # Increase size of x-axis text (specific)
    axis.text.y = element_text(size = 16)   # Increase size of y-axis text (specific)
  )
TotalArea_plot

ggsave("Analysis_Fall2024/Plots/TotalArea_plot.png", plot = TotalArea_plot, height = 6, width = 12, dpi = 300)

```

We can take the code from just above for TA and simply swap out the metric of choice for whatever we want. 

Here I do this for NND and take care to change the y axis labels too.

```{r}
siberDensityPlot(cbind(layman.B[[1]][ , "NND"], 
                       layman.B[[2]][ , "NND"],
                       layman.B[[3]][ , "NND"], 
                       layman.B[[4]][ , "NND"],
                       layman.B[[5]][ , "NND"], 
                       layman.B[[6]][ , "NND"]),
                 xticklabels = c("EYS", "PCH","STB", "SWN","WCW", "WGN"), 
                 bty="L", ylim = c(0, 4),
                 las = 1,
                 ylab = "NND (nearest neighbour distance",
                 xlab = "")

NND_means <- c(mean(layman.B[[1]][ , "NND"]), mean(layman.B[[2]][ , "NND"]), mean(layman.B[[3]][ , "NND"]), 
                  mean(layman.B[[4]][ , "NND"]), mean(layman.B[[5]][ , "NND"]), mean(layman.B[[6]][ , "NND"]))
NND_sds <- c(sd(layman.B[[1]][ , "NND"]), sd(layman.B[[2]][ , "NND"]), sd(layman.B[[3]][ , "NND"]), 
                sd(layman.B[[4]][ , "NND"]), sd(layman.B[[5]][ , "NND"]), sd(layman.B[[6]][ , "NND"]))

NearestNeighbor <- data_frame(names, Status, NND_means, NND_sds)

#Trying to make these plots in ggplot instead as violin plots

# Load required package
library(ggplot2)
library(dplyr)

# Combine the data into a single data frame
layman_data <- do.call(rbind, lapply(seq_along(layman.B), function(i) {
  data.frame(Group = paste0("Group_", i), NND = layman.B[[i]][, "NND"])
}))

# Define the group names for labeling
group_labels <- c("EYS", "PCH", "STB", "SWN", "WCW", "WGN")

# Calculate the 2.5th and 97.5th percentiles for each group
layman_data_filtered <- layman_data %>%
  group_by(Group) %>%
  mutate(
    lower_cutoff = quantile(NND, 0.025),
    upper_cutoff = quantile(NND, 0.975)
  ) %>%
  filter(NND >= lower_cutoff & NND <= upper_cutoff) %>%
  ungroup()

# Calculate medians for each group
median_values <- layman_data_filtered %>%
  group_by(Group) %>%
  summarize(median_NND = median(NND), .groups = "drop")

# Create the violin plot
NND_plot <- ggplot(layman_data_filtered, aes(x = factor(Group, labels = group_labels), y = NND, fill = Group)) +
  geom_violin(trim = TRUE) +
  geom_point(data = median_values, aes(x = factor(Group, labels = group_labels), y = median_NND), 
             shape = 4, size = 3, color = "black", stroke = 1.2) + # Add X for median
  ylab("Nearest Neighbor Distance") +
  xlab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 14),     # Increase size of axis tick labels
    axis.title = element_text(size = 20),   # Increase size of axis titles
    axis.text.x = element_text(size = 16),  # Increase size of x-axis text (specific)
    axis.text.y = element_text(size = 16)   # Increase size of y-axis text (specific)
  )
NND_plot

ggsave("Analysis_Fall2024/Plots/NearestNeighbor_plot.png", plot = NND_plot, height = 6, width = 12, dpi = 300)
```

Here I do this for CD and take care to change the y axis labels too.

```{r}
siberDensityPlot(cbind(layman.B[[1]][ , "CD"], 
                       layman.B[[2]][ , "CD"],
                       layman.B[[3]][ , "CD"], 
                       layman.B[[4]][ , "CD"],
                       layman.B[[5]][ , "CD"], 
                       layman.B[[6]][ , "CD"]),
                 xticklabels = c("EYS", "PCH","STB", "SWN","WCW", "WGN"), 
                 bty="L", ylim = c(0, 5),
                 las = 1,
                 ylab = "CD (centroid distance)",
                 xlab = "")

CD_means <- c(mean(layman.B[[1]][ , "CD"]), mean(layman.B[[2]][ , "CD"]), mean(layman.B[[3]][ , "CD"]), 
                   mean(layman.B[[4]][ , "CD"]), mean(layman.B[[5]][ , "CD"]), mean(layman.B[[6]][ , "CD"]))
CD_sds <- c(sd(layman.B[[1]][ , "CD"]), sd(layman.B[[2]][ , "CD"]), sd(layman.B[[3]][ , "CD"]), 
                 sd(layman.B[[4]][ , "CD"]), sd(layman.B[[5]][ , "CD"]), sd(layman.B[[6]][ , "CD"]))

CentroidDistance <- data_frame(names, Status, CD_means, CD_sds)

# Load required package
library(ggplot2)
library(dplyr)

# Combine the data into a single data frame
layman_data <- do.call(rbind, lapply(seq_along(layman.B), function(i) {
  data.frame(Group = paste0("Group_", i), CD = layman.B[[i]][, "CD"])
}))

# Define the group names for labeling
group_labels <- c("EYS", "PCH", "STB", "SWN", "WCW", "WGN")

# Calculate the 2.5th and 97.5th percentiles for each group
layman_data_filtered <- layman_data %>%
  group_by(Group) %>%
  mutate(
    lower_cutoff = quantile(CD, 0.025),
    upper_cutoff = quantile(CD, 0.975)
  ) %>%
  filter(CD >= lower_cutoff & CD <= upper_cutoff) %>%
  ungroup()

# Calculate medians for each group
median_values <- layman_data_filtered %>%
  group_by(Group) %>%
  summarize(median_CD = median(CD), .groups = "drop")

# Create the violin plot
CD_plot <- ggplot(layman_data_filtered, aes(x = factor(Group, labels = group_labels), y = CD, fill = Group)) +
  geom_violin(trim = TRUE) +
  geom_point(data = median_values, aes(x = factor(Group, labels = group_labels), y = median_CD), 
             shape = 4, size = 3, color = "black", stroke = 1.2) + # Add X for median
  ylab("Centroid Distance") +
  xlab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 14),     # Increase size of axis tick labels
    axis.title = element_text(size = 20),   # Increase size of axis titles
    axis.text.x = element_text(size = 16),  # Increase size of x-axis text (specific)
    axis.text.y = element_text(size = 16)   # Increase size of y-axis text (specific)
  )
CD_plot

ggsave("Analysis_Fall2024/Plots/Centroid_plot.png", plot = CD_plot, height = 6, width = 12, dpi = 300)
```

Here I do this for carbon_range and take care to change the y axis labels too.

```{r}
siberDensityPlot(cbind(layman.B[[1]][ , "dX_range"], 
                       layman.B[[2]][ , "dX_range"],
                       layman.B[[3]][ , "dX_range"], 
                       layman.B[[4]][ , "dX_range"],
                       layman.B[[5]][ , "dX_range"], 
                       layman.B[[6]][ , "dX_range"]),
                 xticklabels = c("EYS", "PCH","STB", "SWN","WCW", "WGN"), 
                 bty="L", ylim = c(0, 20),
                 las = 1,
                 ylab = "Carbon Range",
                 xlab = "")

dCrange_means <- c(mean(layman.B[[1]][ , "dX_range"]), mean(layman.B[[2]][ , "dX_range"]), mean(layman.B[[3]][ , "dX_range"]), 
                  mean(layman.B[[4]][ , "dX_range"]), mean(layman.B[[5]][ , "dX_range"]), mean(layman.B[[6]][ , "dX_range"]))
dCrange_sds <- c(sd(layman.B[[1]][ , "dX_range"]), sd(layman.B[[2]][ , "dX_range"]), sd(layman.B[[3]][ , "dX_range"]), 
                sd(layman.B[[4]][ , "dX_range"]), sd(layman.B[[5]][ , "dX_range"]), sd(layman.B[[6]][ , "dX_range"]))

CarbonRange <- data_frame(names, Status, dCrange_means, dCrange_sds)

# Load required package
library(ggplot2)
library(dplyr)

# Combine the data into a single data frame
layman_data <- do.call(rbind, lapply(seq_along(layman.B), function(i) {
  data.frame(Group = paste0("Group_", i), dX_range = layman.B[[i]][, "dX_range"])
}))

# Define the group names for labeling
group_labels <- c("EYS", "PCH", "STB", "SWN", "WCW", "WGN")

# Calculate the 2.5th and 97.5th percentiles for each group
layman_data_filtered <- layman_data %>%
  group_by(Group) %>%
  mutate(
    lower_cutoff = quantile(dX_range, 0.025),
    upper_cutoff = quantile(dX_range, 0.975)
  ) %>%
  filter(dX_range >= lower_cutoff & dX_range <= upper_cutoff) %>%
  ungroup()

# Calculate medians for each group
median_values <- layman_data_filtered %>%
  group_by(Group) %>%
  summarize(median_dX_range = median(dX_range), .groups = "drop")

# Create the violin plot
dX_range_plot <- ggplot(layman_data_filtered, aes(x = factor(Group, labels = group_labels), y = dX_range, fill = Group)) +
  geom_violin(trim = TRUE) +
  geom_point(data = median_values, aes(x = factor(Group, labels = group_labels), y = median_dX_range), 
             shape = 4, size = 3, color = "black", stroke = 1.2) + # Add X for median
  ylab("Carbon Range") +
  xlab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 14),     # Increase size of axis tick labels
    axis.title = element_text(size = 20),   # Increase size of axis titles
    axis.text.x = element_text(size = 16),  # Increase size of x-axis text (specific)
    axis.text.y = element_text(size = 16)   # Increase size of y-axis text (specific)
  )
dX_range_plot

ggsave("Analysis_Fall2024/Plots/CarbonRange_plot.png", plot = dX_range_plot, height = 6, width = 12, dpi = 300)
```

Here I do this for nitrogen_range and take care to change the y axis labels too.

```{r}
siberDensityPlot(cbind(layman.B[[1]][ , "dY_range"], 
                       layman.B[[2]][ , "dY_range"],
                       layman.B[[3]][ , "dY_range"], 
                       layman.B[[4]][ , "dY_range"],
                       layman.B[[5]][ , "dY_range"], 
                       layman.B[[6]][ , "dY_range"]),
                 xticklabels = c("EYS", "PCH","STB", "SWN","WCW", "WGN"), 
                 bty="L", ylim = c(0, 10),
                 las = 1,
                 ylab = "Nitrogen Range",
                 xlab = "")

dNrange_means <- c(mean(layman.B[[1]][ , "dY_range"]), mean(layman.B[[2]][ , "dY_range"]), mean(layman.B[[3]][ , "dY_range"]), 
                       mean(layman.B[[4]][ , "dY_range"]), mean(layman.B[[5]][ , "dY_range"]), mean(layman.B[[6]][ , "dY_range"]))
dNrange_sds <- c(sd(layman.B[[1]][ , "dY_range"]), sd(layman.B[[2]][ , "dY_range"]), sd(layman.B[[3]][ , "dY_range"]), 
                     sd(layman.B[[4]][ , "dY_range"]), sd(layman.B[[5]][ , "dY_range"]), sd(layman.B[[6]][ , "dY_range"]))

NitrogenRange <- data_frame(names, Status, dNrange_means, dNrange_sds)

#Trying to make these plots in ggplot instead as violin plots

# Load required package
library(ggplot2)
library(dplyr)

# Combine the data into a single data frame
layman_data <- do.call(rbind, lapply(seq_along(layman.B), function(i) {
  data.frame(Group = paste0("Group_", i), dY_range = layman.B[[i]][, "dY_range"])
}))

# Define the group names for labeling
group_labels <- c("EYS", "PCH", "STB", "SWN", "WCW", "WGN")

# Calculate the 2.5th and 97.5th percentiles for each group
layman_data_filtered <- layman_data %>%
  group_by(Group) %>%
  mutate(
    lower_cutoff = quantile(dY_range, 0.025),
    upper_cutoff = quantile(dY_range, 0.975)
  ) %>%
  filter(dY_range >= lower_cutoff & dY_range <= upper_cutoff) %>%
  ungroup()

# Calculate medians for each group
median_values <- layman_data_filtered %>%
  group_by(Group) %>%
  summarize(median_dY_range = median(dY_range), .groups = "drop")

# Create the violin plot
dY_range_plot <- ggplot(layman_data_filtered, aes(x = factor(Group, labels = group_labels), y = dY_range, fill = Group)) +
  geom_violin(trim = TRUE) +
  geom_point(data = median_values, aes(x = factor(Group, labels = group_labels), y = median_dY_range), 
             shape = 4, size = 3, color = "black", stroke = 1.2) + # Add X for median
  ylab("Nitrogen Range") +
  xlab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 14),     # Increase size of axis tick labels
    axis.title = element_text(size = 20),   # Increase size of axis titles
    axis.text.x = element_text(size = 16),  # Increase size of x-axis text (specific)
    axis.text.y = element_text(size = 16)   # Increase size of y-axis text (specific)
  )
dY_range_plot

ggsave("Analysis_Fall2024/Plots/NitrogenRange_plot.png", plot = dY_range_plot, height = 6, width = 12, dpi = 300)
```

Making the Layman Metrics plots in ggplot

```{r}
library(ggplot2)

ggplot(TotalArea, aes(x = names, y = TA_means, fill = Status))+
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar(aes(x = names, ymin = TA_means - TA_sds, ymax = TA_means + TA_sds), width = 0.2, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Pond", y = "Total Area")+
  ggtitle("Comparing Total Isotopic Area")+
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 15), legend.title = element_text(size = 20))

ggplot(NearestNeighbor, aes(x = names, y = NND_means, fill = Status))+
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar(aes(x = names, ymin = NND_means - NND_sds, ymax = NND_means + NND_sds), width = 0.2, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Pond", y = "Nearest Neighbor Distance")+
  ggtitle("Comparing Nearest Neighbor Distance")+
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 15), legend.title = element_text(size = 20))


ggplot(CentroidDistance, aes(x = names, y = CD_means, fill = Status))+
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar(aes(x = names, ymin = CD_means - CD_sds, ymax = CD_means + CD_sds), width = 0.2, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Pond", y = "Centroid Distance")+
  ggtitle("Comparing Centroid Distance")+
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 15), legend.title = element_text(size = 20))


ggplot(CarbonRange, aes(x = names, y = dCrange_means, fill = Status))+
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar(aes(x = names, ymin = dCrange_means - dCrange_sds, ymax = dCrange_means + dCrange_sds), width = 0.2, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Pond", y = "Carbon Range")+
  ggtitle("Comparing Carbon Range")+
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 15), legend.title = element_text(size = 20))


ggplot(NitrogenRange, aes(x = names, y = dNrange_means, fill = Status))+
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar(aes(x = names, ymin = dNrange_means - dNrange_sds, ymax = dNrange_means + dNrange_sds), width = 0.2, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Pond", y = "Nitrogen Range")+
  ggtitle("Comparing Nitrogen Range")+
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), plot.title = element_text(size = 22, hjust = 0.5),
        legend.text = element_text(size = 15), legend.title = element_text(size = 20))
```

Running t-tests to compare the total isotopic area etc. between elodea and no elodea lakes

```{r}
t.test(TA_means ~ Status, data = TotalArea)

t.test(NND_means ~ Status, data = NearestNeighbor)

t.test(CD_means ~ Status, data = CentroidDistance)

t.test(dCrange_means ~ Status, data = CarbonRange)

t.test(dNrange_means ~ Status, data = NitrogenRange)
```