---
title: "T-tests Macrophytes Invertebrates"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting up the workspace

```{r}
library(SIBER, quietly = TRUE,
        verbose = FALSE,
        logical.return = FALSE)
library(viridis)
library(tibble)
library(dplyr)
```

# Comparisons of C and N signatures Based on Elodea Presence/Absence

Starting with Macrophytes

Read in the data
```{r}
mydata <- read.csv("~/Dropbox/SWEL/CopperRiverDelta/Elodea/SI_Ponds/DataAnalysis/ElodeaStableIsotopes/Analysis_Fall2024/DataFiles/CRD_MacrophyteData_SIBER.csv")
CRD_data <- mydata
```

Add a status column here
```{r}
CRD_data

CRD_data <- CRD_data %>%
  mutate(status = case_when(
    community %in% c("EYS", "PCH", "STB") ~ "Native",
    community %in% c("SWN", "WCW", "WGN") ~ "Elodea",
    TRUE ~ NA_character_  # Handle any other cases
  ))
CRD_data
```

Comparing the isotopic signatures of macrophytes based on Elodea Presence/Absence

## Bogbean

```{r}
Bogbean <- CRD_data %>%
  filter(group == "Bogbean")

ttest_carbon <- t.test(iso1 ~ status, data = Bogbean, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Bogbean, na.action = na.omit)
print(ttest_nitrogen)
```
## Horsetail
```{r}
Horsetail <- CRD_data %>%
  filter(group == "Horsetail")

ttest_carbon <- t.test(iso1 ~ status, data = Horsetail, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Horsetail, na.action = na.omit)
print(ttest_nitrogen)
```

## Millfoil
```{r}
Millfoil <- CRD_data %>%
  filter(group == "Millfoil")

ttest_carbon <- t.test(iso1 ~ status, data = Millfoil, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Millfoil, na.action = na.omit)
print(ttest_nitrogen)
```

## Peat Moss
```{r}
PeatMoss <- CRD_data %>%
  filter(group == "Peat Moss")

ttest_carbon <- t.test(iso1 ~ status, data = PeatMoss, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = PeatMoss, na.action = na.omit)
print(ttest_nitrogen)
```

## Sedge
```{r}
Sedge <- CRD_data %>%
  filter(group == "Sedge")

ttest_carbon <- t.test(iso1 ~ status, data = Sedge, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Sedge, na.action = na.omit)
print(ttest_nitrogen)
```

## Yellow Pondlily
```{r}
YellowPondlily <- CRD_data %>%
  filter(group == "Yellow Pondlily")

ttest_carbon <- t.test(iso1 ~ status, data = YellowPondlily, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = YellowPondlily, na.action = na.omit)
print(ttest_nitrogen)
```

## Bladderwort
```{r}
#Bladderwort <- CRD_data %>%
 # filter(group == "Bladderwort")

#ttest_carbon <- t.test(iso1 ~ status, data = Bladderwort, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = Bladderwort, na.action = na.omit)
#print(ttest_nitrogen)
```

## Cotton Grass
```{r}
CottonGrass <- CRD_data %>%
  filter(group == "Cotton Grass")

ttest_carbon <- t.test(iso1 ~ status, data = CottonGrass, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = CottonGrass, na.action = na.omit)
print(ttest_nitrogen)
```

## Purple Marshlocks
```{r}
PurpleMarshlocks <- CRD_data %>%
  filter(group == "Purple Marshlocks")

ttest_carbon <- t.test(iso1 ~ status, data = PurpleMarshlocks, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = PurpleMarshlocks, na.action = na.omit)
print(ttest_nitrogen)
```

## Mare's Tail
```{r}
MaresTail <- CRD_data %>%
  filter(group == "Mare's Tail")

ttest_carbon <- t.test(iso1 ~ status, data = MaresTail, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = MaresTail, na.action = na.omit)
print(ttest_nitrogen)
```

## Pondweed
```{r}
Pondweed <- CRD_data %>%
  filter(group == "Pondweed")

ttest_carbon <- t.test(iso1 ~ status, data = Pondweed, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Pondweed, na.action = na.omit)
print(ttest_nitrogen)
```

## Algae
```{r}
#Algae <- CRD_data %>%
 # filter(group == "Algae")

#ttest_carbon <- t.test(iso1 ~ status, data = Algae, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = Algae, na.action = na.omit)
#print(ttest_nitrogen)
```

## Burreed
```{r}
#Burreed <- CRD_data %>%
 # filter(group == "Burreed")

#ttest_carbon <- t.test(iso1 ~ status, data = Burreed, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = Burreed, na.action = na.omit)
#print(ttest_nitrogen)
```

## Reed Grass
```{r}
#ReedGrass <- CRD_data %>%
 # filter(group == "Reed Grass")

#ttest_carbon <- t.test(iso1 ~ status, data = ReedGrass, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = ReedGrass, na.action = na.omit)
#print(ttest_nitrogen)
```

# Continuing with the Invertebrates

Read in the data
```{r}
mydata <- read.csv("~/Dropbox/SWEL/CopperRiverDelta/Elodea/SI_Ponds/DataAnalysis/ElodeaStableIsotopes/Analysis_Fall2024/DataFiles/CRD_SIData.csv",
                   header=TRUE)
```

Add a status column here
```{r}
mydata

mydata <- mydata %>%
  mutate(status = case_when(
    group %in% c("EyakSouth", "ThePinch", "Strawberry") ~ "Native",
    group %in% c("Swanny", "WestCanneryWest", "Wigeon") ~ "Elodea",
    TRUE ~ NA_character_  # Handle any other cases
  ))
mydata
```

## Amphipod
```{r}
Amphipod <- mydata %>%
  filter(community == "Amphipod")

ttest_carbon <- t.test(iso1 ~ status, data = Amphipod, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Amphipod, na.action = na.omit)
print(ttest_nitrogen)
```

## Beetle
```{r}
Beetle <- mydata %>%
  filter(community == "Beetle")

ttest_carbon <- t.test(iso1 ~ status, data = Beetle, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Beetle, na.action = na.omit)
print(ttest_nitrogen)
```

## Caddisfly
```{r}
Caddisfly <- mydata %>%
  filter(community == "Caddisfly")

ttest_carbon <- t.test(iso1 ~ status, data = Caddisfly, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Caddisfly, na.action = na.omit)
print(ttest_nitrogen)
```

## Clams
```{r}
Clams <- mydata %>%
  filter(community == "Clams")

ttest_carbon <- t.test(iso1 ~ status, data = Clams, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Clams, na.action = na.omit)
print(ttest_nitrogen)
```

## Damselfly
```{r}
Damselfly <- mydata %>%
  filter(community == "Damselfly")

ttest_carbon <- t.test(iso1 ~ status, data = Damselfly, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Damselfly, na.action = na.omit)
print(ttest_nitrogen)
```


## Dragonfly
```{r}
Dragonfly <- mydata %>%
  filter(community == "Dragonfly")

ttest_carbon <- t.test(iso1 ~ status, data = Dragonfly, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Dragonfly, na.action = na.omit)
print(ttest_nitrogen)
```

## Fly
```{r}
Fly <- mydata %>%
  filter(community == "Fly")

ttest_carbon <- t.test(iso1 ~ status, data = Fly, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Fly, na.action = na.omit)
print(ttest_nitrogen)
```

## Leech
```{r}
#Leech <- mydata %>%
 # filter(community == "Leech")

#ttest_carbon <- t.test(iso1 ~ status, data = Leech, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = Leech, na.action = na.omit)
#print(ttest_nitrogen)
```

## Mite
```{r}
Mite <- mydata %>%
  filter(community == "Mite")

ttest_carbon <- t.test(iso1 ~ status, data = Mite, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = Mite, na.action = na.omit)
print(ttest_nitrogen)
```

## Snail
```{r}
#Snail <- mydata %>%
 # filter(community == "Snail")

#ttest_carbon <- t.test(iso1 ~ status, data = Snail, na.action = na.omit)
#print(ttest_carbon)

#ttest_nitrogen <- t.test(iso2 ~ status, data = Snail, na.action = na.omit)
#print(ttest_nitrogen)
```

## Water Boatman
```{r}
WaterBoatman <- mydata %>%
  filter(community == "Water Boatman")

ttest_carbon <- t.test(iso1 ~ status, data = WaterBoatman, na.action = na.omit)
print(ttest_carbon)

ttest_nitrogen <- t.test(iso2 ~ status, data = WaterBoatman, na.action = na.omit)
print(ttest_nitrogen)
```

